---
title: "R Notebook"
output: html_notebook
---

The aim of analyzing the provided crime and temperature data is to gain insights into the relationship between various factors such as weather conditions and crime incidents.Explore patterns and trends in crime incidents.This will help identify areas with higher crime rates and understand the nature of incidents that occur.Provide recommendations for crime prevention: Based on our analysis, we aim to identify potential strategies or interventions that can help mitigate the impact of weather-related factors on crime.

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)
library(e1071)
library(mlbench)
library(ggplot2)
library(gridExtra)

```

We have loaded the data of the temperature in which all the date , temperature , wind and its speed , pressure at sea level in hectopascals (hPa) , total cloud cover in octas , low cloud cover in octas , sunshine during in hour of that day , Visibility in kilometers, Snow depth in centimeters these are few column



```{r}
df<- read.csv("temp2023.csv")
head(df)
```




```{r}
df$Date <- as.Date(df$Date, format="%m/%d/%Y")
df
df_avg <- df %>%
  group_by(month = format(Date, "%m"), year = format(Date, "%Y")) %>%
  summarise(
    across(
      where(is.numeric),
      mean,
      na.rm = TRUE
    ),
    .groups = 'drop'
  ) %>%
  mutate(date_formatted = paste(year, month, sep="-"))

df_avg$date_formatted


df_avg%>%head(10)

```
```{r}
df1<- read.csv("crime23.csv")

df1%>%head(10)

```



```{r}
merged_data <- merge(df_avg, df1, by.x="date_formatted", by.y="date", all=TRUE)
merged_data%>%head(10)
```


```{r}
names(merged_data)
```



```{r}
merged_data %>%
select(date_formatted) %>%
count(date_formatted, sort = TRUE, name = "Count")%>%
ggplot(aes(factor(date_formatted), Count, group = 1)) +     
geom_bar(stat = "identity", fill = "red", color = "black") +
geom_line(color = "black") +
geom_smooth() +
geom_point(color = "green") +
geom_text(aes(label = Count, vjust = -0.3), color = "black") +
scale_x_discrete(guide = guide_axis(angle = 90)) +
labs(x = "Month", y = "count of  Incident categeory") +
theme_minimal()
```




```{r}
summary(merged_data)
```




```{r}
merged_data %>%
  select(category) %>%
  count(category, sort = TRUE, name = "Count")
```

```{r}
# Create a subset of merged_data containing only "street_name" and "category" columns
subset_data <- merged_data[, c("date_formatted", "category")]

# Remove rows with missing values
subset_data <- na.omit(subset_data)

# Plot a bar graph
barplot(table(subset_data$date_formatted, subset_data$category), 
        main = "Category Distribution by date_formatted",
        xlab = "category",
        ylab = "Count",
        las = 2, # Rotate x-axis labels vertically
        col = rainbow(length(unique(subset_data$category))),  # Assign different colors to each category
        legend = TRUE)  # Show legend


```

```{r}
# Calculate the count and percentage for each category
d <- merged_data %>%
  count(category) %>%
  mutate(perc = round(100 * n / sum(n)))

# Plot the pie chart
ggplot(data = d, aes(x = "", y = n, fill = category)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(perc, "%")), position = position_stack(vjust = 0.5)) +
  labs(fill = 'Category', x = NULL, y = NULL, title = 'Incidents by Category', subtitle = 'Percentage of Total') +
  coord_polar(theta = "y") +
  theme_minimal()

```


```{r}
# Remove rows with NA values
cleaned_data <- na.omit(merged_data[, c("TemperatureCAvg", "TemperatureCMin", "TemperatureCMax", "SnowDepcm")])

# Dot plot with custom color palette and summary statistics
p <- ggplot(cleaned_data, aes(x = SnowDepcm, color = "TemperatureCAvg")) +
  geom_point(aes(y = TemperatureCAvg), size = 3) +
  geom_point(aes(y = TemperatureCMin, color = "TemperatureCMin"), size = 3) +
  geom_point(aes(y = TemperatureCMax, color = "TemperatureCMax"), size = 3) +
  scale_color_manual(values = c("TemperatureCAvg" = "#999999", "TemperatureCMin" = "#E69F00", "TemperatureCMax" = "#56B4E9")) +
  labs(x = "Snow Depth (cm)", y = "Temperature (°C)", color = "Variables") +
  theme_minimal() +
  stat_summary(aes(y = TemperatureCAvg), fun = mean, geom = "point", shape = 18, size = 3, color = "red") +  # Add mean points
  stat_summary(aes(y = TemperatureCAvg), fun = median, geom = "point", shape = 18, size = 3, color = "blue")  # Add median points

# Preview the plot with custom color palette and summary statistics
p

```

```{r}
print(colnames(merged_data))
```

```{r}
library(patchwork)

# Histogram for WindkmhInt
hist_windkmhInt <- ggplot(merged_data, aes(x = WindkmhInt)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.6) +
  labs(title = "WindSpeed(km/h)-Int", x = "Wind Speed (km/h)", y = "Frequency")

# Histogram for WindkmhGust
hist_windkmhGust <- ggplot(merged_data, aes(x = WindkmhGust)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black", alpha = 0.6) +
  labs(title = "Wind Speed(km/h)-Gust", x = "Wind Speed (km/h)", y = "Frequency")

# Histogram for VisKm
hist_VisKm <- ggplot(merged_data, aes(x = VisKm)) +
  geom_histogram(binwidth = 1, fill = "salmon", color = "black", alpha = 0.6) +
  labs(title = "Visibility(km)", x = "Visibility (km)", y = "Frequency")

# Arrange histograms together
hist_windkmhInt + hist_windkmhGust + hist_VisKm + plot_layout(ncol = 3)

```



```{r}
merged_data <- merged_data[!is.na(merged_data$outcome_status), ]
summary(merged_data$outcome_status)
```

```{r}


```


```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a scatter plot with jittered points
sinaplot <- ggplot(merged_data, aes(x = street_id, y = category)) +
  geom_jitter(width = 0.3, height = 0.3) +  # Jitter the points to avoid overlap
  labs(x = "street_id", y = "Category", title = "Sinaplot of Outcome Status and Category")

# Print the sinaplot
print(sinaplot)

```
```{r}
library(ggplot2)

# Create a jitter plot
ggplot(merged_data, aes(x = date_formatted, y = category)) +
  geom_jitter(alpha = 0.5) +
  labs(x = "Date Formatted", y = "category", title = "Relation between Date Formatted and Street Name") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # Rotate x-axis labels

```

```{r}
# Calculate the correlation matrix
correlation_matrix <- cor(merged_data[, c("TemperatureCAvg", "TemperatureCMax", "TemperatureCMin", "SnowDepcm")])

# Visualize the correlation matrix using a heatmap
library(ggplot2)

# Convert the correlation matrix to a data frame
correlation_df <- as.data.frame(as.table(correlation_matrix))

# Plot the heatmap
ggplot(correlation_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_fixed()


```



```{r}
print(colnames(merged_data))
```


  
```{r}

```


```{r}
# Load necessary libraries
library(plotly)
library(dplyr)

# Filter out rows with missing street names or categories
filtered_data <- merged_data %>%
  filter(!is.na(street_name) & !is.na(category))

# Create scatter plot
scatter_plot <- plot_ly(filtered_data, x = ~street_name, y = ~category, type = 'scatter', mode = 'markers') %>%
  layout(title = 'Scatter Plot of Categories by Street Names',
         xaxis = list(title = 'Street Name'),
         yaxis = list(title = 'Category'),
         hovermode = 'closest')

# Print the scatter plot
scatter_plot
```


```{r}
# Load necessary libraries
library(plotly)
library(dplyr)

# Filter out rows with missing street names or categories
filtered_data <- merged_data %>%
  filter(!is.na(street_name) & !is.na(category))

# Aggregate data to count occurrences of each category by street name
count_data <- filtered_data %>%
  group_by(street_name, category) %>%
  summarise(count = n()) %>%
  ungroup()

# Create scatter plot
scatter_plot <- plot_ly(count_data, x = ~street_name, y = ~category, z = ~count,
                        type = 'scatter3d', mode = 'markers', text = ~paste("Count: ", count)) %>%
  layout(title = 'Scatter Plot of Categories by Street Names',
         scene = list(xaxis = list(title = 'Street Name'),
                      yaxis = list(title = 'Category'),
                      zaxis = list(title = 'Count')),
         hovermode = 'closest')

# Print the scatter plot
scatter_plot

```



```{r}
# Time series plot
ggplot(category_counts, aes(x = date_formatted, y = count, color = category, group = category)) +
  geom_line() +
  labs(title = "Crime Trends in Colchester (2023)",
       x = "Date",
       y = "Number of Crimes",
       color = "Category") +
  theme_minimal()

```
```{r}
# Time series plot with smoothing
ggplot(category_counts, aes(x = date_formatted, y = count, color = category, group = category)) +
  geom_line() +
  geom_smooth(method = "auto", se = FALSE) +  # Add smoothing line
  labs(title = "Crime Trends in Colchester (2023)",
       x = "Date",
       y = "Number of Crimes",
       color = "Category") +
  theme_minimal()

```

```{r}
colnames(merged_data) <- paste0("col", 1:5) 
rownames(merged_data) <- paste0("row", 1:5) 
  
# Draw a heatmap 
heatmap(merged_data)         
```

